---
title: "AR1 model using log ER"
author:
date:
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# knitr::opts_knit$set(root.dir = 'C:/Users/Alice Carter/git/autotrophic_rivers/')
source('C:/Users/Alice Carter/git/autotrophic_rivers/src/stan_helpers.R')
library(rstan)
```

Goal: To build an AR1 model of log(ER)

I'm not completely confident in how I have made the transformation to a log model, input would be welcomed.

### Process model:

$$log(R_t) = \alpha_0 + \phi log(R_{t-1}) + w_t$$
$$0 < \phi < 1$$
$$w_t \sim normal(0, \sigma_{proc})$$

Which can be rewritten as:
$$R_t = e^{\alpha_0} R_{t-1}^{\phi} e^ {w_t}$$
or,
$$R_t = \beta_0R_{t-1}^\phi \epsilon_t$$
$$\epsilon_t \sim lognormal(0, \sigma_{proc})$$

### Observation model:

$$R_{obs} = R * \epsilon_t$$
$$\epsilon_t \sim lognormal(0, \sigma_{obs})$$

or, 

$$log(R_{obs}) = log(R) + w_t$$
$$w_t \sim normal(0, \sigma_{obs})$$

### Stan Code

```{r, eval = FALSE}

sink('ar1_model_logR.stan')
cat("
  data {
    int <lower = 0> N;  // Sample size
    vector[N] logR_obs; // log ER observed
    real mu_obs;        // prior mean for obs error
  }
  
  parameters {
    real b0; // Intercept
    real <lower = 0, upper = 1> phi; 
    real <lower = 0> sigma_proc;
    real <lower = 0> sigma_obs;
    vector <lower = 0> [N] logR; // State variable for ER
  }
  
  model {
    logR[1] ~ normal(logR_obs[1], sigma_obs);
    logR[2:N] ~ normal(b0 + phi * logR[1:(N-1)], sigma_proc);
    logR_obs ~ normal(logR, sigma_obs);
   
    // priors
    b0 ~ normal(0,1);
    phi ~ beta(1,1);
    sigma_proc ~ normal(0,1) T[0,];
    sigma_obs ~ normal(mu_obs, mu_obs/2) T[0,];
  }
  
  generated quantities {
    vector [N] R_hat;   // estimated underlying state
    vector [N] R_tilde; // estimated log ER obs
    R_hat[1] = logR[1];
    R_tilde[1] = logR[1];
    for(i in 2:N){
      R_hat[i] = normal_rng(b0 + phi * R_hat[i-1], sigma_proc);
      R_tilde[i] = normal_rng(R_hat[i], sigma_obs);  
    }
  }", fill = T)

sink()

```

### Simulated Data

Simulate data with an expected value of ER = 5,

$log(ER) = log(5) = \frac{\beta_0}{1-\phi}$

set $\beta_0 = 1$

$log(5) = \frac{1}{1-\phi}$

$\phi = 0.38$

```{r}
b0 = 1      # intercept
phi = 0.38  # ar1 coefficient
sigma_proc = 0.1 # ~10% error
sigma_obs = 0.05  

logR <- numeric(100)  # ER state variable
logR[1] <- log(5)

# Observed ER, on a log scale
logR_obs = rnorm(1, logR[1], sigma_obs)
for (i in 2:100){
  logR[i] <- b0 + phi * logR[i-1] + rnorm(1, 0, sigma_proc)
  logR_obs[i] = rnorm(1, logR[i], sigma_obs)
}

# compare state variable and observations
par(mfrow = c(1,2))

plot(exp(logR), type = 'l', ylab = 'ER', xlab = 'days')
points(exp(logR_obs))

# does variance scale with ER?
plot(exp(logR), abs(exp(logR_obs) - exp(logR)), xlab = 'ER', ylab = 'obs error')

```

### Run Model

```{r, eval = F}
sim_dat <- list(logR_obs = logR_obs, N = length(logR_obs), mu_obs = sigma_obs)
fit <- stan(file = 'src/SAM/stan/ar1_model.stan', 
            data = sim_dat,
            warmup = 500, iter = 1000, 
            chains = 4, cores = 4)

```

### Parameter recovery
```{r}
fit <- readRDS('C:/Users/Alice Carter/git/autotrophic_rivers/src/ar1/log_ar1_sim_fit.rds')
traceplot(fit, pars= c('b0', 'phi', 'sigma_proc', 'sigma_obs'))
plot_post_sim(fit, 
              pars = c('b0', 'phi', 'sigma_proc', 'sigma_obs'),
              vals = c(b0, phi, sigma_proc, sigma_obs))
```

Here I put a fairly strong prior on observation error,

$\sigma_{obs} \sim Normal(0.05, 0.025)$

but I am still getting poorly mixed chains and the observation error is highly correlated with the log posterior
```{r}
pairs(fit, pars = c('b0', 'phi', 'sigma_proc', 'sigma_obs', 'lp__'))
```
