---
title: "AR1 model using log ER"
author: "alice carter"
date: "11/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/Alice Carter/git/autotrophic_rivers/')
library(rstan)
```

Problem: the uncertainty in estimates of ER scale with the magnitude of ER

Solution: model the logarithm of ER

Issues: I get errors when I rewrite the stan ar1 model so that the observation error is incorporated on the log of respiration, where $R$ is observed respiration and $mu$ is the underlying state variable as:

$$ log(R) \sim Normal(log(mu), \sigma_{obs})$$

Here I have included the revised stan code and simulated data I am using to test it as well as the errors I am getting when I try to run the model.

### Stan Code

```{r, eval = FALSE}

sink('src/SAM/stan/ar1_model.stan')
cat("
  data {
    int <lower = 0> N; // Sample size
    vector[N] logR;    // log ER observed
    real mu_obs;       // prior mean for obs error
  }
  
  parameters {
    real a0; // Intercept
    real <lower = 0, upper = 1> a1; 
    real <lower = 0> sigma_proc;
    real <lower = 0> sigma_obs;
    vector <lower = 0> [N] mu; // State variable for ER
  }
  
  model {
    mu[1] ~ normal(exp(logR[1]), 0.1);
    mu[2:N] ~ normal(a0 + a1 * mu[1:(N-1)], sigma_proc);
    logR ~ normal(log(mu), sigma_obs);
   
    // priors
    a0 ~ normal(0,1);
    a1 ~ uniform(0,1);
    sigma_proc ~ normal(0,1) T[0,];
    sigma_obs ~ normal(mu_obs, mu_obs/2) T[0,];
  }
  
  generated quantities {
    vector [N] R_hat;   // estimated underlying state
    vector [N] R_tilde; // estimated log ER obs
    R_hat[1] = exp(logR[1]);
    R_tilde[1] = logR[1];
    for(i in 2:N){
      R_hat[i] = normal_rng(a0 + a1 * R_hat[i-1], sigma_proc);
      R_tilde[i] = normal_rng(log(R_hat[i]), sigma_obs);  
    }
  }", fill = T)

sink()

```

### Simulated Data

```{r}
a0 = 1    # intercept
a1 = 0.8  # ar1 coefficient
sigma_proc = 0.2
sigma_obs = 0.1

mu <- numeric(100)  # ER state variable
mu[1] <- 5

# Observed ER, on a log scale
R = log(mu[1]) + rnorm(1, 0, sigma_obs)
for (i in 2:100){
  mu[i] <- a0 + a1 * mu[i-1] + rnorm(1, 0, sigma_proc)
  R[i] = log(mu[i]) + rnorm(1, 0, 0.2)
}

# compare state variable and observations
par(mfrow = c(1,2))

plot(mu, type = 'l')
points(exp(R))

# does variance scale with ER?
plot(mu, abs(exp(R) - mu), xlab = 'ER', ylab = 'obs error')

```

### Run Model

```{r}
sim_dat <- list(logR = R, N = length(R), mu_obs = sigma_obs)
fit <- stan(file = 'src/SAM/stan/ar1_model.stan', data = sim_dat,
            warmup = 500, iter = 1000, 
            chains = 4, cores = 4)

```



The error suggests running with only one chain to debug. Doing that gives the following errors:

```{r}
fit <- stan(file = 'src/SAM/stan/ar1_model.stan', data = sim_dat,
            warmup = 500, iter = 1000, 
            chains = 1, cores = 4)
```

It looks like there is a problem with my generated quatnities, but I am not sure why there is a problem with the way it is specified. Taking the generated quantities out of the model and rerunning it gives me this:

### Model without generated quantities block

```{r}
fit <- stan(file = 'src/SAM/stan/ar1_model_nogq.stan', data = sim_dat,
            warmup = 500, iter = 1000, 
            chains = 4, cores = 4)
```
I'm still getting a lot of warnings, and I am not doing a great job recovering my parameters (true values shown below as red triangles). Additionally, the posteriors for a0 and a1 are very highly correlated. Any advice about how I should be changing the model to incorporate log(ER) would be helpful.
```{r}
p <- rstan::plot(fit, show_density = T, fill_color = 'grey',
                 pars = c('a0', 'a1', 'sigma_obs', 'sigma_proc'))
dd <- data.frame(x = c(1, 0.8, 0.1, 0.2), y = c(4,3,2,1))
p + geom_point(data = dd, aes(x = x, y = y), size = 3, shape = 17, col = 'brown3')

pairs(fit, pars = c('a0', 'a1', 'sigma_obs', 'sigma_proc'))
